---
title: "Airlines and Flights datasets analysis"
author: "Oscar Hernandez Mata"
date: "2022"
output: html_document
---

In this project I will be working with R, SQL, and Python in the same document. I will use the data sets *airlines* and *flights* from the package *nycflights13*.

```{r, warning=FALSE, message=FALSE}

#install.packages("nycflights13") 
##install.packages("sqldf")       
#install.packages("tidyverse")    
#install.packages("dplyr")


library(nycflights13)
library(sqldf)
library(tidyverse)
library(dplyr)
library(haven)
library(ggplot2)
library(data.table)

data(nycflights13::airlines)
data(nycflights13::airports)
data(nycflights13::flights)
data(nycflights13::planes)
data(nycflights13::weather)

data("airlines")
data("airports")
data("flights")
data("planes")
data("weather")
```


> **I listed the name of airlines where the destination is ATL airport with their *average arrival delays* and sort them from the smallest to largest average arrival delays.** 

I solved this using R, SQL, and Python.

### R solution

```{r, warning=FALSE, message=FALSE}

atlanta_flights <- filter(flights, dest == "ATL")
joined_atlanta_airlines <- inner_join(atlanta_flights, airlines, by = "carrier")

avg_delays <- joined_atlanta_airlines %>%
  group_by(name) %>%
  summarize(average_arrival_delay = mean(arr_delay, na.rm = TRUE)) %>%
  arrange(average_arrival_delay)

avg_delays

```


### SQL solution

```{r, warning=FALSE, message=FALSE }
sqldf("
  SELECT airlines.name, AVG(arr_delay) AS average_arrival_delay
  FROM flights
  JOIN airlines ON flights.carrier = airlines.carrier
  WHERE dest = 'ATL'
  GROUP BY airlines.name
  ORDER BY average_arrival_delay ASC
")
```


### Python solution

```{python}

import pandas as pd
import numpy as np
import statsmodels.api as stat

from nycflights13 import airlines, flights

flights_atlanta = flights[flights['dest'] == 'ATL']

data0 = flights_atlanta.join(airlines.set_index('carrier'), on='carrier')

data_average_delay = data0.groupby('name')['arr_delay'].mean().reset_index()

data_average_delay = data_average_delay.sort_values('arr_delay')

data_average_delay

```



> **I ploted the boxplot of the departure delays vs the name of airlines where the destination is ATL airport. I solved this using R and Python.**


### R solution
Use ggplot
```{r, warning=FALSE, message=FALSE}

atl_flights <- filter(flights, dest == "ATL")

ggplot(joined_atlanta_airlines, aes(x = name, y = dep_delay)) + 
  geom_boxplot() +
  ggtitle("Departure Delays by Airline to ATL Airport") +
  xlab("Airline Name") + ylab("Departure Delay (minutes)")


```


### Python solution

```{python}
import pandas as pd
import matplotlib.pyplot as plt

from nycflights13 import flights, airlines

atl_flights = flights[flights['dest'] == "ATL"]

data0.boxplot(column='dep_delay', by='name')
plt.title("Departure Delays by Airline to ATL")
plt.xlabel("Airline")
plt.ylabel("Departure Delay (minutes)")
plt.show()

```

> **For each airlines: I found the month where the average departure delay time is the highest in the year. I Made a visualization to show the results. I solved this question using Python.**

```{python}

grouped = data0.groupby('month')['dep_delay'].mean()
grouped

fig, ax = plt.subplots(figsize=(10, 6))
ax.bar(grouped.index, grouped.values)
ax.set_xticks(range(1, 13))
ax.set_xlabel('Month')
ax.set_ylabel('Average departure delay (min)')
ax.set_title('Average departure delay time by month across all airlines')
plt.show()
```
The month withe highest average departure delay is July. 
